---
title: "IUU_Index_Analysis"
author: "Chloe Swick"
date: "2023-12-21"
output: html_document
---

```{r setup, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(dplyr)
library(janitor)
library(readxl)
library(countrycode)
library(broom)

```



# Loading engagment metrics per organization 

```{r}
organization_engagements <- read_csv("Skylight Program Metrics_Organizations_Table.csv")
```


# Loading engagement metrics per country 

```{r}
weekly_country_engagements <- read_csv("Skylight Program Metrics_Countries_Table.csv")
```



# Loading total engagements per country and organization 

```{r}
skylight_countries <- read_csv("Metrics_Countries_Pivot_table-2.csv")
```


# Loading IUU index indicator scores 2019-2023 

```{r}
indicator_scores <- read_csv("iuu_fishing_index_2019-2023_indicator_scores.csv")
```


# Loading Skylight deal dates 

```{r}
skylight_deals <- read_excel("updated deals.xlsx")
```



# Cleaning data 


```{r}
skylight_deals <- clean_names(skylight_deals)
iuu_index_scores <- clean_names(indicator_scores)
organization_engagements <- clean_names(organization_engagements)
country_engagements_total <- clean_names(skylight_countries)
weekly_country_engagements <- clean_names(weekly_country_engagements)
```


```{r}
weekly_country_engagements$country <- countrycode(weekly_country_engagements$country, "country.name", "country.name")

iuu_index_scores$country[iuu_index_scores$country == "Micronesia (FS of)"] <- "Federated States of Micronesia"
iuu_index_scores$country <- countrycode(iuu_index_scores$country, "country.name", "country.name")

skylight_deals$deal_country[skylight_deals$deal_country == "Micronesia"] <- "Federated States of Micronesia"

skylight_deals$country <- countrycode(skylight_deals$deal_country, "country.name", "country.name") 
```



# Compiling data 

join by country 

```{r}
colnames(country_engagements_total)[1] = "country"
```

```{r}
iuu_index_countries <- c(unique(iuu_index_scores$country))

#view(iuu_index_scores)

skylight_countries <- c(unique(weekly_country_engagements$country))

#view(skylight_countries)

```



```{r}

# Get unique values from both vectors
all_values <- unique(c(iuu_index_countries, skylight_countries))

# Create a data frame with NAs
common_countries_df <- data.frame(country = all_values, iuu_index_countries = NA, skylight_countries = NA)

# Fill in values in the data frame based on the presence in vectors
common_countries_df$iuu_index_countries[match(iuu_index_countries, common_countries_df$country)] <- iuu_index_countries
common_countries_df$skylight_countries[match(skylight_countries, common_countries_df$country)] <- skylight_countries

common_countries_df$treatment_yes_no <- ifelse(is.na(common_countries_df$skylight_countries), 0, 1)


# Print the result
#print(common_countries_df) ## good start but need to figure out a good way to get rid of the spelling differences (micronesia and south korea). maybe we can ask Chris
```


```{r}
#weekly_country_engagements$active_yes_no <- ifelse(weekly_country_engagements$no_of_active_weeks>4, 1, 0)
```



```{r}
dind_df <- iuu_index_scores %>% 
  full_join(common_countries_df, by = join_by(country)) %>%
  full_join(weekly_country_engagements, by = join_by(country))
```


```{r}

ctk <- c("deal_title", "country", "deal_deployment_date", "deal_organisation","organization_type")

active_country_dates_df <- skylight_deals %>% 
  subset(select = ctk) %>% 
  filter(deal_deployment_date > 2019) %>% 
  group_by(country) %>%
  filter(deal_deployment_date == min(deal_deployment_date, na.rm = TRUE)) %>%
  ungroup() %>%  #filter out duplicate country entries for country level analysis 
  distinct(country, deal_deployment_date, .keep_all = TRUE)

```



```{r}
active_country_dates_clean <- active_country_dates_df %>% 
  mutate(treatment_year = as.integer(format(as.Date(deal_deployment_date), "%Y"))) %>% 
  filter(organization_type != "Regional") %>% #come back to this assumption later
  select(country, treatment_year) %>% 
  rename(deal_country = country)
```


```{r}
dind_id_1 <- dind_df %>% 
  filter(indicator_id == 1)
```



```{r}
dind_df_1 <- dind_id_1 %>% 
  left_join(active_country_dates_clean %>% 
    select(deal_country, treatment_year), 
    by = c("skylight_countries" = "deal_country")) %>% 
  rename(skylight_treatment = treatment_yes_no) %>% 
  rename(index_year = year) %>% 
  filter(!(skylight_treatment == 1 & is.na(treatment_year))) %>% #removing skylight countries without deal date 
  mutate(pre_post_treatment = case_when(
    skylight_treatment == 0 & index_year < 2021 ~ 0,
    skylight_treatment == 0 & index_year >= 2021 ~ 1,
    skylight_treatment == 1 & index_year < treatment_year ~ 0,
    skylight_treatment == 1 & index_year >= treatment_year ~ 1,
    TRUE ~ NA_integer_  # Default case, you can change NA to another value if needed
  )) 
### still want to add the number of sessions as a measure of engagement



```


Trying with all of the indicators 

```{r}
dind_df_all <- dind_df %>% 
  left_join(active_country_dates_clean %>% 
    select(deal_country, treatment_year), 
    by = c("skylight_countries" = "deal_country")) %>% 
  rename(skylight_treatment = treatment_yes_no) %>% 
  rename(index_year = year) %>% 
  filter(!(skylight_treatment == 1 & is.na(treatment_year))) %>% #removing skylight countries without deal date 
  mutate(pre_post_treatment = case_when(
    skylight_treatment == 0 & index_year < 2021 ~ 0,
    skylight_treatment == 0 & index_year >= 2021 ~ 1,
    skylight_treatment == 1 & index_year < treatment_year ~ 0,
    skylight_treatment == 1 & index_year >= treatment_year ~ 1,
    TRUE ~ NA_integer_  # Default case, you can change NA to another value if needed
  )) 
```


# Try the regression!


regression for loop! 

```{r}
results_by_indicator <- data.frame()

for (i in 1:40) {
  # Subset the data for each indicator value
  subset_data <- dind_df_all[dind_df_all$indicator_id == i, ]
  
  # Run linear regression with interaction
  model <- lm(score ~ skylight_treatment + pre_post_treatment + skylight_treatment * pre_post_treatment, data = subset_data)
  
 tidy_results <- tidy(model) %>%
    mutate(indicator = i)
  
  # Bind the results to the main data frame
  results_by_indicator <- bind_rows(results_by_indicator, tidy_results)
}

# Print the results (optional)
print(results_by_indicator)


# Export to CSV file (optional)
write.csv(results_by_indicator, file = "linear_model_results.csv", row.names = FALSE)

```

```{r}
significant_p <- results_by_indicator %>% 
  filter(p.value <= .05)

sig_p_pre_post <- significant_p %>% 
  filter(term == "pre_post_treatment")
```



```{r}
dind_analysis_1 <- lm(score ~ skylight_treatment + pre_post_treatment + skylight_treatment * pre_post_treatment, data = dind_df_1)

summary(dind_analysis_1)
```



# Data Visualizations

figure out how many NA values there are per indicator 

```{r}
ggplot(dind_df_final, aes(x = year, y = score, color = skylight_treatment, group = interaction(indicator_id, skylight_treatment))) +
  stat_summary(fun = "mean", geom = "line", size = 1) +
  labs(title = "Mean Score Over Years by Skylight Treatment",
       x = "Year",
       y = "Mean Score",
       color = "Skylight Treatment") +
  facet_wrap(~ indicator_id) +
  theme_minimal()
```



```{r}
ggplot(subset(dind_df_final, indicator_id == "32"), aes(x = year, y = score,group = country, color = skylight_treatment, group = interaction(indicator_id, skylight_treatment))) +
  geom_point(position = position_jitter(width = 0.2, height = 0), alpha = 0.5)+
  theme(legend.position = "none")
```








