---
title: "iuuf_indicator_analysis"
author: "E.M.Thomas"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(dplyr)
library(janitor)
library(readxl)
library(countrycode)
library(broom)
library(here)
library(tidyr)
library(jtools)
library(caret)
library(AICcmodavg) 
library(equatiomatic)
```

### Load data 
```{r}
organization_engagements <- read_csv(here("data", "Skylight Program Metrics_Organizations_Table.csv"))
weekly_country_engagements <- read_csv(here("data", "Skylight Program Metrics_Countries_Table.csv"))
skylight_engagement_sessions <- read_csv(here("data","Metrics_Countries_Pivot_table-2.csv"))
indicator_scores <- read_csv(here("data","iuu_fishing_index_2019-2023_indicator_scores.csv"))
skylight_deals <- read_excel(here("data","updated deals.xlsx"))
governance_indicators <- read_csv(here("data", "governance_indicators.csv"))
#fish_indicators <- read_csv(here("data", "WDIData.csv"))
fish_id_wide <- read_csv(here("data", "fish_indicators.csv"))
wb_gdp_raw <- read_csv(here("data", "World Bank GDP.csv"))
```

### Clean data
```{r}

###clean names 
skylight_deals <- clean_names(skylight_deals)
iuu_index_scores <- clean_names(indicator_scores)
organization_engagements <- clean_names(organization_engagements)
#country_engagements_total <- clean_names(skylight_countries)
weekly_country_engagements <- clean_names(weekly_country_engagements)
governance_indicators <- clean_names(governance_indicators) %>% 
  rename(country=country_name)
skylight_engagement_sessions <- clean_names(skylight_engagement_sessions) 
#fish_indicators_clean <- clean_names(fish_indicators)
wb_gdp_clean <- wb_gdp_raw %>% clean_names() %>% 
  rename(country=country_name)

weekly_country_engagements$country <- countrycode(weekly_country_engagements$country, "country.name", "country.name") 

weekly_country_engagements$no_of_active_weeks <- as.numeric(weekly_country_engagements$no_of_active_weeks)

#filter for only engaged countries
weekly_country_engagements <- weekly_country_engagements%>% 
  filter(no_of_active_weeks >2) 

#list of active skylight countries 
active_skylight_countries <- weekly_country_engagements$country 

iuu_index_scores$country[iuu_index_scores$country == "Micronesia (FS of)"] <- "Federated States of Micronesia"
iuu_index_scores$country <- countrycode(iuu_index_scores$country, "country.name", "country.name")


skylight_deals$deal_country[skylight_deals$deal_country == "Micronesia"] <- "Federated States of Micronesia"

skylight_deals$country <- countrycode(skylight_deals$deal_country, "country.name", "country.name") 

country_sessions <- skylight_engagement_sessions %>% 
  rename(country = engagement)

country_sessions$country <- countrycode(country_sessions$country, "country.name", "country.name") 

country_sessions <- country_sessions %>% 
  drop_na(country) %>% 
  group_by(country) %>%
  summarize(sessions = sum(sessions, na.rm = TRUE))


###IUUFR Data

#colnames(country_engagements_total)[1] = "country"
iuu_index_countries <- c(unique(iuu_index_scores$country))
#skylight_countries <- c(unique(weekly_country_engagements$country))

# Get unique values from both vectors
all_values <- unique(c(iuu_index_countries, active_skylight_countries))

# Create a data frame with NAs
common_countries_df <- data.frame(country = all_values, iuu_index_countries = NA, active_skylight_countries = NA)

# Fill in values in the data frame based on the presence in vectors
common_countries_df$iuu_index_countries[match(iuu_index_countries, common_countries_df$country)] <- iuu_index_countries
common_countries_df$active_skylight_countries[match(active_skylight_countries, common_countries_df$country)] <- active_skylight_countries

common_countries_df$skylight_treatment <- ifelse(is.na(common_countries_df$active_skylight_countries), 0, 1)




#####cleaning governance indicator data 

governance_indicators$country <- countrycode(governance_indicators$country,"country.name", "country.name")

governance_indicators <- governance_indicators %>%
  rename_all(~ gsub("x(\\d+)_yr\\1", "\\1", .))

gov_id_estimates <- governance_indicators %>% 
  filter(grepl("EST", series_code)) %>% 
  mutate_all(~ ifelse(. == "..", NA, .)) %>%
  na.omit()

gov_id_long <- gov_id_estimates %>%
  pivot_longer(cols = 5:28,
               names_to = "year", 
               values_to = "value") %>% 
  subset(select = -series_code) %>% 
  pivot_wider(names_from = "series_name",
              values_from = "value") %>% 
  clean_names()

### cleaning gdp data
wb_gdp_clean$country <- countrycode(wb_gdp_clean$country, "country.name", "country.name")
wb_gdp_df <- wb_gdp_clean %>% 
  rename_all(~ gsub("x(\\d+)_yr\\1", "\\1", .)) %>% 
  select(-series_name, -series_code, -country_code)

wb_gdp_long <- wb_gdp_df %>% 
  pivot_longer(cols = 2:13,
               names_to = "year", 
               values_to = "gdp") %>% 
  mutate_all(~ ifelse(. == "..", NA, .)) %>%
  na.omit() %>% 
  filter(year > 2017)

wb_gdp_long$gdp <- as.numeric(wb_gdp_long$gdp)


wb_gdp_sum <- wb_gdp_long %>% 
  group_by(country) %>% 
  summarise_all(mean, na.rm=TRUE) %>% 
  select(-year)  %>% 
  filter(country %in% iuu_index_countries)


####cleaning fishing indicator data 

#fish_indicators_clean <- fish_indicators_clean %>% 
 # filter(indicator_name %in% c("Aquaculture production (metric tons)", "Capture fisheries production (metric tons)", "Fish species, threatened", "Total fisheries production (metric tons)")) 
  
#ish_indicators_clean$country <- countrycode(fish_indicators_clean$country_name,"country.name", "country.name")


```


dont run this 
#```{r}
fish_indicators_clean <- fish_indicators_clean %>% 
  filter(country %in% iuu_index_countries) %>% 
  select(-x1960:-x2017) %>% 
  select(-x2022, -x68)

fish_indicators_clean <- fish_indicators_clean %>%
  rename_all(~ gsub("x(\\d+)", "\\1", .))

fish_id_long <- fish_indicators_clean %>% 
  pivot_longer(cols = 5:30,
               names_to = "year",
               values_to = "value")

fish_id_wide <- fish_id_long %>% 
  subset(select = -indicator_code) %>% 
  pivot_wider(names_from = "indicator_name", 
              values_from = "value") %>% 
  clean_names() %>% 
   subset(select = -fish_species_threatened) 
#```

##downloading the data:
#```{r}
write.csv(fish_id_wide, "/Users/chloe/Desktop/IUU_Analysis/IUU_Index_Analysis/data/fish_indicators.csv", row.names=FALSE)
#```



```{r}

iuu_index_df <- iuu_index_scores %>% 
  full_join(common_countries_df, by = join_by(country)) %>%
  full_join(weekly_country_engagements, by = join_by(country))

ctk <- c("deal_title", "country", "deal_deployment_date", "deal_organisation","organization_type")

active_country_dates_df <- skylight_deals %>% 
  subset(select = ctk) %>% 
  filter(deal_deployment_date > 2019) %>% 
  group_by(country) %>%
  filter(deal_deployment_date == min(deal_deployment_date, na.rm = TRUE)) %>%
  ungroup() %>%  #filter out duplicate country entries for country level analysis 
  distinct(country, deal_deployment_date, .keep_all = TRUE)

active_country_dates_clean <- active_country_dates_df %>% 
  mutate(treatment_year = as.integer(format(as.Date(deal_deployment_date), "%Y"))) %>% 
  filter(organization_type != "Regional") %>% #come back to this assumption later
  select(country, treatment_year) #%>% 
 # rename(deal_country = country)

iuu_index_final_df <- iuu_index_df %>% 
  left_join(active_country_dates_clean, 
            by = c("active_skylight_countries" = "country")) %>% 
  rename(index_year = year) %>% 
  filter(!(skylight_treatment == 1 & is.na(treatment_year))) #removing skylight countries without deal date 

iuu_index_final_df <- iuu_index_final_df %>% 
  left_join(country_sessions %>% 
  select(country, sessions), 
  by = "country")
```


### Compile WB Governance data
```{r}
wb_gov_index_df <- gov_id_long %>% 
  filter(country %in% iuu_index_countries) %>% 
  full_join(common_countries_df %>% 
      select(country, skylight_treatment),
    by = join_by(country)) %>%
  full_join(weekly_country_engagements,
    by = join_by(country)) %>% 
  full_join(country_sessions,
    by = join_by(country)) %>% 
  mutate(
    no_of_active_weeks = ifelse(is.na(no_of_active_weeks), 0, no_of_active_weeks),
    sessions = ifelse(is.na(sessions), 0, sessions)) %>% 
  filter(year > 2017) %>% 
  select(!country_code) %>% 
  na.omit()
  

wb_gov_index_df$year <- as.numeric(wb_gov_index_df$year)
wb_gov_index_df$control_of_corruption_estimate <- as.numeric(wb_gov_index_df$control_of_corruption_estimate)
wb_gov_index_df$government_effectiveness_estimate <- as.numeric(wb_gov_index_df$government_effectiveness_estimate)
wb_gov_index_df$political_stability_and_absence_of_violence_terrorism_estimate <- as.numeric(wb_gov_index_df$political_stability_and_absence_of_violence_terrorism_estimate)
wb_gov_index_df$regulatory_quality_estimate <- as.numeric(wb_gov_index_df$regulatory_quality_estimate)
wb_gov_index_df$rule_of_law_estimate <- as.numeric(wb_gov_index_df$rule_of_law_estimate)
wb_gov_index_df$voice_and_accountability_estimate <- as.numeric(wb_gov_index_df$voice_and_accountability_estimate)
```


### Compile Fishing Idicator Data 
```{r}
fishing_skylight_df <- fish_id_wide %>% 
  full_join(common_countries_df %>% 
              select(country, skylight_treatment),
            by = join_by(country)) %>% 
    full_join(weekly_country_engagements,
    by = join_by(country)) %>% 
  full_join(country_sessions,
    by = join_by(country)) %>% 
  mutate(
    no_of_active_weeks = ifelse(is.na(no_of_active_weeks), 0, no_of_active_weeks),
    sessions = ifelse(is.na(sessions), 0, sessions)) %>% 
  na.omit() %>% 
  filter(year > 2017)
  
```

```{r}
#str(fishing_skylight_df)
fishing_skylight_df$skylight_treatment <- as.factor(fishing_skylight_df$skylight_treatment )
```



```{r}
# combine gov and fishing indicators
all_index_df <- fishing_skylight_df %>% 
  select(country, year, aquaculture_production_metric_tons, capture_fisheries_production_metric_tons, total_fisheries_production_metric_tons) %>%
  full_join(wb_gov_index_df, by= join_by(year, country)) %>% 
  #group_by(country) %>% 
  #summarise_all(mean,na.rm = TRUE) %>% 
  full_join(wb_gdp_sum, join_by(country)) %>% 
  filter(country != "Montenegro")

all_index_df$ratio_capture <- all_index_df$aquaculture_production_metric_tons/all_index_df$capture_fisheries_production_metric_tons

all_index_df$skylight_treatment <- as.factor(all_index_df$skylight_treatment)
```


```{r}
##compiling IUU index 
iuu_final_final <- iuu_index_final_df %>% 
  filter(index_year < "2023") %>% 
  select(country, indicator_id, score, index_year, skylight_treatment, no_of_active_weeks) %>% 
 pivot_wider(names_from = "indicator_id", 
             values_from = "score") %>% 
    select_if(~ !any(is.na(.))) %>% #removes indicators that have NA values b/c regression no likey 
  #group_by(country) %>%
  #select(-country, index_year, skylight_treatment) %>%  # Exclude non-numeric column if necessary
  #summarise_all(mean, na.rm = TRUE) %>% 
  select(-index_year) %>% 
  clean_names()
  
```

```{r}

iuu_final_final$skylight_treatment <- as.factor(iuu_final_final$skylight_treatment )
#str(iuu_final_final)
```

### Linear Models
we should split up analysis by nonprofit and government, do we see a difference? 

Skylight Analysis 1: skylight treatment = aquaculture production + capture fishery 
  *** want to add ratio of capture fishery to aquaculutre and total percent of GDP that comes from fishing 
  
```{r}
skylight_iuu_1 <- glm(skylight_treatment ~ x29 + x3 + x4 + x5 + x10 + x12 + x14, data = iuu_final_final, family = binomial)

summary(skylight_iuu_1)
```


```{r}
skylight_iuu_2 <- glm(skylight_treatment ~ x4 + x10 + x14, data = iuu_final_final, family = binomial)

summary(skylight_iuu_2)
```

  
```{r}
skylight_analysis_1 <- glm(skylight_treatment ~ aquaculture_production_metric_tons + capture_fisheries_production_metric_tons, data = fishing_skylight_df, family = binomial)

summary(skylight_analysis_1)
```

#there is a way to interpret results in terms of odds, we could talk about this in presentation 

```{r}
skylight_analysis_2 <- glm(skylight_treatment ~ aquaculture_production_metric_tons + capture_fisheries_production_metric_tons, data = all_index_df, family = binomial)

summary(skylight_analysis_2)
```


# GLM 

```{r} 
#w/o gdp
f1 = skylight_treatment ~ 
                      aquaculture_production_metric_tons +
                      capture_fisheries_production_metric_tons +
                      control_of_corruption_estimate +
                      government_effectiveness_estimate +
                      political_stability_and_absence_of_violence_terrorism_estimate+
                      regulatory_quality_estimate +
                      rule_of_law_estimate +
                       ratio_capture +
                      voice_and_accountability_estimate

skylight_glm1 <- glm(f1, data = all_index_df, family = binomial)

#w/gdp
f2 = skylight_treatment ~ 
                      aquaculture_production_metric_tons +
                      capture_fisheries_production_metric_tons +
                      control_of_corruption_estimate +
                      government_effectiveness_estimate +
                      political_stability_and_absence_of_violence_terrorism_estimate+
                      regulatory_quality_estimate +
                      rule_of_law_estimate + gdp +
                      voice_and_accountability_estimate

skylight_glm2 <- glm(f2, data = all_index_df, family = binomial)

#reduced to significant & significant ish 
f3 = skylight_treatment ~ 
                      control_of_corruption_estimate +
                      regulatory_quality_estimate +
                       ratio_capture+
                      voice_and_accountability_estimate

skylight_glm3 <- glm(f3,data = all_index_df, family = binomial)

aictab(list(skylight_glm1, skylight_glm2, skylight_glm3)) #model 3 is best

bictab(list(skylight_glm1, skylight_glm2, skylight_glm3)) #model 3 is best

```


```{r}
summary(skylight_glm3)
summary(skylight_glm2)
summary(skylight_glm1)
```


```{r}
skylight_analysis_final <- glm(skylight_treatment ~ control_of_corruption_estimate + ratio_capture + voice_and_accountability_estimate, data = all_index_df, family = binomial)

summary(skylight_analysis_final)
```



```{r}
prediction <- predict(skylight_analysis_final, all_index_df)

# Ensure that both prediction and actual values are factors with the same levels
prediction <- factor(ifelse(prediction > 0.5, "1", "0"))
actual <- all_index_df$skylight_treatment

# Create confusion matrix
confusion_matrix <- table(Predicted = prediction, Actual = actual)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

# Display confusion matrix and accuracy
print(confusion_matrix)
cat("Accuracy:", accuracy, "\n")

```


```{r}
#confusionMatrix(data=skylight_glm1, all_index_df)
```



```{r}
plot_final1 <- ggplot(data = all_index_df, aes(x=skylight_treatment, y=control_of_corruption_estimate)) +
  geom_boxplot(aes(color = skylight_treatment)) +
  scale_color_manual(values = c("0" = "#336e94ff", "1" = "#0097a7ff")) +
  labs(x = "Skylight Use (0 = no, 1 = yes)", y = "Control of Corruption Estimate") +
  theme_minimal()

plot_final2 <- ggplot(data = all_index_df, aes(x=skylight_treatment, y=ratio_capture)) +
  geom_boxplot(aes(color = skylight_treatment))+
  scale_color_manual(values = c("0" = "#336e94ff", "1" = "#0097a7ff"))+
   labs(x = "Skylight Use (0 = no, 1 = yes)", y = "Aquaculture/Capture Production") +
  theme_minimal()

plot_final3 <- ggplot(data = all_index_df, aes(x=skylight_treatment, y=voice_and_accountability_estimate)) +
  geom_boxplot(aes(color = skylight_treatment))+
  scale_color_manual(values = c("0" = "#336e94ff", "1" = "#0097a7ff"))+
   labs(x = "Skylight Use (0 = no, 1 = yes)", y = "Voice and Accountability Estimates") +
  theme_minimal()

plot_final1 
plot_final2
plot_final3

ggsave(plot_final1 , filename = "corruption_box.png", bg = "transparent")
ggsave(plot_final2, filename = "fishing_box.png", bg = "transparent")
ggsave(plot_final3, filename = "accountability_box.png", bg = "transparent")

```


```{r}
ggplot(data = all_index_df, aes(x=skylight_treatment, y=ratio_capture)) +
  geom_violin(aes(color = skylight_treatment))

```

```{r}
ggplot(data = all_index_df, aes(x=skylight_treatment, y=control_of_corruption_estimate)) +
  geom_violin(aes(color = skylight_treatment))

```

```{r}
plot_box3 <- ggplot(data = all_index_df, aes(x=skylight_treatment, y=voice_and_accountability_estimate)) +
  geom_violin(aes(color = skylight_treatment))

plot_box3
```


```{r}
plot_box3 <- ggplot(data = all_index_df, aes(x=skylight_treatment, y=gdp)) +
  geom_boxplot(aes(color = skylight_treatment))

plot_box3
```


```{r}
#volume of catches
plot_box4 <- ggplot(data = iuu_final_final, aes(x=skylight_treatment, y=x29)) +
  geom_boxplot(aes(color = skylight_treatment))

plot_box4
```


```{r}
#EEZ size 
plot_box5 <- ggplot(data = iuu_final_final, aes(x=skylight_treatment, y=x10)) +
  geom_boxplot(aes(color = skylight_treatment))

plot_box5
```




```{r}
#EEZ size violin 
ggplot(data = iuu_final_final, aes(x=skylight_treatment, y=x10)) +
  geom_violin(aes(color = skylight_treatment))


```


```{r}
#MSC certified fisheries 
plot_box5 <- ggplot(data = iuu_final_final, aes(x=skylight_treatment, y=x14)) +
  geom_boxplot(aes(color = skylight_treatment))

plot_box5
```

```{r}
ggplot(data = all_index_df, aes(x=skylight_treatment, y = capture_fisheries_production_metric_tons)) +
  geom_violin(aes(color = skylight_treatment))
```

```{r}
y_prob <- predict(skylight_analysis_final, newdata = all_index_df, type = 'response')
all_index_df$y_prob <- y_prob

ggplot(all_index_df, aes(x=control_of_corruption_estimate, y=skylight_treatment)) + geom_point() +
     geom_smooth(method = "glm", color = "green", se = TRUE, 
              method.args = list(family = "binomial"), formula = y ~ poly(x, 2))
### need to hold other variables constant, can only graph one at a time, pick the best one? 
```


```{r}
# Load the pROC package
install.packages("pROC")
library(pROC)

# Get predicted probabilities
y_prob <- predict(skylight_analysis_final, newdata = all_index_df, type = 'response')

# Create a ROC curve object
roc_curve <- roc(all_index_df$skylight_treatment, y_prob)

# Calculate AUC
auc_value <- auc(roc_curve)

# Print the AUC value
print(auc_value)
```

```{r}
y_prob <- predict(skylight_analysis_final, newdata = all_index_df, type = 'response')

# Create a ROC curve object
roc_curve <- roc(all_index_df$skylight_treatment, y_prob)

# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)

# Add diagonal reference line for random guessing
abline(a = 0, b = 1, col = "red", lty = 2)

# Calculate and print AUC
auc_value <- auc(roc_curve)
print(paste("AUC:", auc_value))
```



```{r}
effect_plot(skylight_glm3,
            voice_and_accountability_estimate,
            interval = TRUE,
            y.label = "Probability")
```


### we can look at a variable that is the ratio of aquaculture to capture fisheries 
I also wonder if we could find data on the total percentage of GDP that comes from capture fisheries

### trying to make a visualization 
```{r}
# Assuming skylight_analysis_1 is your logistic regression model
library(ggplot2)

# Create a sequence of values for the predictor variable
predictor_values <- seq(min(fishing_skylight_df$predictor), max(fishing_skylight_df$predictor), length.out = 100)

# Predict probabilities using the model
predicted_probs <- predict(skylight_analysis_1, newdata = data.frame(predictor = predictor_values), type = "response")

# Combine predictor values and predicted probabilities into a data frame
plot_data <- data.frame(predictor = predictor_values, predicted_prob = predicted_probs)

# Create a plot
ggplot(fishing_skylight_df, aes(x = predictor, y = skylight_treatment)) +
  geom_point() +  # Scatter plot of the data points
  geom_line(data = plot_data, aes(x = predictor, y = predicted_prob), color = "blue") +  # Logistic regression line
  theme_minimal() +
  labs(title = "Logistic Regression Plot", x = "Predictor Variable", y = "Probability of skylight treatment")

```





